---
output: github_document
---

# read-rs-product

How to read common remote sensing products.

## Table of content

- [L3BIN](#l3bin)
- [NetCDF4](#netcdf-4)

## L3BIN

Applies to:
- MODIS Terra and Aqua

HDF5 L3BIN files can be read using the `rhdf5` package from bioconductor. The installation can be done as follow:

```{r, eval = FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite("rhdf5")
```

The `h5ls()` function will list all the layers included in the file:

```{r}
library(rhdf5)

h5ls("data/A2016160.L3b_DAY_CHL.nc")
```

Finally, use `h5read()` to open a specific layer:

```{r}
df <- h5read("data/A2016160.L3b_DAY_CHL.nc", "/level-3_binned_data/chlor_a")
head(df)
```

Where:

- `sum`  = the sum of the pixels in the bin
- `sum_squared` = the squared sum

It is to be noted that the observed values need to be weighted. The weighted values can be found in the `BinList` layer:

```{r}
bins <- h5read("data/A2016160.L3b_DAY_CHL.nc", "/level-3_binned_data/BinList")

df$sum <- df$sum / bins$weights

head(df)
```

In Matlab:

```matlab
res = h5read('A2016009.L3b_DAY_CHL.nc', '/level-3_binned_data/chlor_a');
```

## NetCDF 4

These files can be opened using the `ncdf4` package.

```{r, eval = FALSE}
install.packages("ncdf4")
```

Layers in the file can be listed using the `nc_open()` function:

```{r}
library(ncdf4)
f <- nc_open("data/avhrr-only-v2.20160503.nc")
f
```

Opening a specific layer is done with the `ncvar_get()` function:

```{r}
sst <- ncvar_get(f, "sst")
dim(sst)
```

It is also possible to open NetCDF4 file using the `raster()` function:

```{r}
library(raster)
r <- raster("data/avhrr-only-v2.20160503.nc", varname = "sst") 
r

plot(r)
```

Note the longitudes are from 0 to 360 degrees. It is easy to change that so the image is correctly displayed:

```{r}
library(sp)

coords <- coordinates(r)
head(coords)

coords[1, ] <- ifelse(coords[1, ] > 180, coords[1, ] - 360, coords[1, ])
# coordinates(r) <- coords

plot(r)
```

## HDF4

http://hdfeos.org/software/r.php

```{r}
library(rgdal)
library(gdalUtils)
sds <- get_subdatasets("data/MYD08_D3.A2003181.051.2008343213114.hdf")

head(sds)

dat <- readGDAL(sds[6])

plot(dat)
```

