---
output: github_document
editor_options:
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk[["set"]](
  comment = "#>",
  echo = TRUE,
  fig.width = 6L
)
```

# read-rs-product

How to read common remote sensing products.

## Table of content

- [L3BIN](#l3bin)
- [NetCDF4](#netcdf-4)

## L3BIN

Applies to:

- MODIS Terra and Aqua

HDF5 L3BIN files can be read using the `rhdf5` package from bioconductor. The installation can be done as follow:

```{r, eval = FALSE}
# https://bioconductor.org/install/

if (!require("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}

BiocManager::install()

BiocManager::install("rhdf5")

library("rhdf5")
```

The `h5ls()` function will list all the layers included in the file:

```{r}
library(rhdf5)

h5ls("data/A2016160.L3b_DAY_CHL.nc")
```

Finally, use `h5read()` to open a specific layer:

```{r}
df <- h5read("data/A2016160.L3b_DAY_CHL.nc", "/level-3_binned_data/chlor_a")
head(df)
```

Where:

- `sum` = the sum of the pixels in the bin
- `sum_squared` = the squared sum

It is to be noted that the observed values need to be weighted. The weighted values can be found in the `BinList` layer:

```{r}
bins <- h5read("data/A2016160.L3b_DAY_CHL.nc", "/level-3_binned_data/BinList")

df[["sum"]] <- df[["sum"]] / bins[["weights"]]

head(df)
```

## NetCDF 4

These files can be opened using the `ncdf4` package.

```{r, eval = FALSE}
install.packages("ncdf4")
```

Layers in the file can be listed using the `nc_open()` function:

```{r}
library(ncdf4)
f <- nc_open("data/avhrr-only-v2.20160503.nc")
f
```

Opening a specific layer is done with the `ncvar_get()` function:

```{r}
sst <- ncvar_get(f, "sst")
dim(sst)
```

It is also possible to open NetCDF4 file using the `rast()` function:

```{r}
library(terra)
r <- rast("data/avhrr-only-v2.20160503.nc", lyrs = 1L)

r

plot(r)
```

Rotate the raster:

```{r}
r <- rotate(r)

plot(r)
```


## HDF4

http://hdfeos.org/software/r.php

```{r terra-hdf, cache=TRUE}
library(terra)

filename <- "data/MYD08_D3.A2003181.051.2008343213114.hdf"

sdss <- describe(filename, sds = TRUE, meta = FALSE, parse = FALSE)

head(sdss)

r <- rast(filename, 142L)

plot(r)
```

## NSIDC data

```{r}
sic <- readBin("data/nsidc/nt_20171002_f17_v1.1_n.bin",
  what = "integer",
  n = 304 * 448,
  size = 1,
  signed = F
)

lat <-
  readBin(
    "data/nsidc/psn25lats_v3.dat",
    what = "integer",
    n = 304 * 448,
    size = 4,
    signed = TRUE
  )

lon <-
  readBin(
    "data/nsidc/psn25lons_v3.dat",
    what = "integer",
    n = 304 * 448,
    size = 4,
    signed = TRUE
  )

df <- data.frame(lat = lat / 100000, lon = lon / 100000, thickness = sic)

plot(df$lon, df$lat)
```

```{r}
library(tidyverse)
df %>%
  ggplot(aes(x = lon, y = lat, color = sic)) +
  geom_point()
```

## Sea ice concentration

ftp://ftp.awi.de/sea_ice/product/amsr2

```{r}
library(terra)

r <- rast("/vsicurl/ftp://ftp.awi.de/sea_ice/product/amsr2/v110/nh/2022/08/nh_SIC-LEADS_2022080400_080412.tiff")

r

plot(r)
```

## h5e file

https://gist.github.com/mdsumner/08265bac3faf1e4325e2d5a39a5f66bf

```{r}
library(terra)
library(gdalraster)
library(stars)

file <- "./data/AMSR_U2_L3_SeaIce12km_B04_20140314.he5"
sub <- paste0("NetCDF:", file, ":")

sds <- gdal_subdatasets(sub)

sf::gdal_utils("info", unlist(sds)[30])

dsn <- "/vsimem/temp.vrt"
sf::gdal_utils(
  "warp",
  unlist(sds)[30],
  dsn,
  options = c("-t_srs", "EPSG:3411", "-overwrite")
)

r <- rast(dsn)

plot(r)
```
