# Binary format

```{r}
#| label: setup
library(terra)
```

## NSIDC binary data

Binary data is a format where information is stored in a sequence of bytes, representing data in its most basic form without any text-based structure. This format is efficient for storage and quick to read programmatically, but it's not human-readable without proper interpretation. In the case of scientific datasets like sea ice concentration, binary formats were often used to minimize file size and optimize data processing speed.

::: {.callout-note}

Version 1 of the [NSIDC-0081 sea ice concentration dataset](https://nsidc.org/data/nsidc-0081/versions/1) is available in binary format but is no longer active. [The current version (Version 2)](https://nsidc.org/data/nsidc-0081/versions/2) is in NetCDF format.

:::

Based on the [documentation](https://nsidc.org/sites/default/files/nsidc-0081-v001-userguide_1.pdf), the binary files can be read using the following code:

```{r}
sic <- readBin(
  here::here("data", "nsidc", "nt_20171002_f17_v1.1_n.bin"),
  what = "integer",
  n = 304L * 448L,
  size = 1L,
  signed = FALSE
)


lat <- readBin(
  here::here("data", "nsidc", "psn25lats_v3.dat"),
  what = "integer",
  n = 304L * 448L,
  size = 4L,
  signed = TRUE
)

lon <- readBin(
  here::here("data", "nsidc", "psn25lons_v3.dat"),
  what = "integer",
  n = 304L * 448L,
  size = 4L,
  signed = TRUE
)
```

We can construct a data frame from the latitude, longitude, and sea ice concentration data:

```{r}
df <- data.frame(lat = lat / 100000L, lon = lon / 100000L, sic = sic)

head(df)

range(df[["sic"]], na.rm = TRUE)
```

According to the documentation:

> The sea ice concentration floating-point values (fractional coverage ranging from 0.0 to 1.0) are
> multiplied by a scaling factor of 250. To convert to the fractional range of 0.0 to 1.0, divide the
> scaled data in the file by 250.

| Data Value | Description                                                                                                                                      |
| ---------- | ------------------------------------------------------------------------------------------------------------------------------------------------ |
| 0 - 250    | Sea ice concentration (fractional coverage scaled by 250)                                                                                        |
| 251        | Circular mask used in the Arctic to cover the irregularly-shaped data gap around the pole (caused by the orbit inclination and instrument swath) |
| 252        | Unused                                                                                                                                           |
| 253        | Coast                                                                                                                                            |
| 254        | Land                                                                                                                                             |
| 255        | Missing data                                                                                                                                     |

### Conversion to raster

We can turn the data into a raster object using the `terra` package. The sea ice concentration values are stored in a 448x304 matrix, which we can convert to a raster object.

```{r}
sic_matrix <- matrix(sic, nrow = 448L, ncol = 304L, byrow = TRUE)

r <- rast(sic_matrix)

plot(r)

ext(r) <- c(-3850000L, 3750000L, -5350000L, 5850000L)
crs(r) <- "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=-45 +k=1 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"

plot(r)
```
