# HE5 format

This section demonstrates how to access and visualize sea ice concentration data stored in HE5 files, focusing on data from the AMSR-E/AMSR2 Unified L3 Daily 12.5 km dataset. We'll use R along with the `gdalraster`, `stars`, `terra`, and `sf` packages.

::: {.callout-note}

## Disclaimer

The code below is largely based/copied on the work of [mdsumner](https://gist.github.com/mdsumner/08265bac3faf1e4325e2d5a39a5f66bf).

:::

```{r}
#| label: setup
library(gdalraster)
library(stars)
library(terra)
```

```{r}
file <- here::here("data", "AMSR_U2_L3_SeaIce12km_B04_20140314.he5")
```

The code leverages the NetCDF interface to access subdatasets within the HE5 file. This allows us to access the geolocation arrays, which is crucial for proper spatial referencing.

```{r}
sub <- paste0("NetCDF:", file, ":")
sds <- gdal_subdatasets(sub)

sf::gdal_utils("info", unlist(sds)[[30L]])
```

The data needs to be transformed to a standard coordinate reference system (CRS). EPSG:3411 (NSIDC Sea Ice Polar Stereographic North) is commonly used for Northern Hemisphere sea ice data. The `gdal_utils()` function warps the data to this CRS and saves it to a temporary VRT (Virtual Raster) file. VRT files are useful because they act as a pointer to the original data without duplicating it, and allow for on-the-fly transformations.

```{r}
dsn <- "/vsimem/temp.vrt"

sf::gdal_utils(
  "warp",
  unlist(sds)[[30L]],
  dsn,
  options = c("-t_srs", "EPSG:3411", "-overwrite")
)
```

The `rast()` function from the `terra` package is used to read the VRT file and create a `SpatRaster` object. This object contains the sea ice concentration data, which can be visualized using the `plot()` function.

```{r}
r <- rast(dsn)

plot(r)
```

## Using `describe()` to List available layers

Simpler, can read the data directly from the HE5 file.

```{r}
var_names <- describe(file, sds = TRUE)

head(var_names)

layer <- var_names[["var"]][grepl("SI_12km_NH_ICECON_DAY", var_names[["var"]], fixed = TRUE)]

layer
```

```{r}
ice_layer <- rast(file, subds = layer)
```

Why are the dimensions different?

```{r}
dim(r)
dim(ice_layer)
```

```{r}
plot(ice_layer)
```
