# GDAL Virtual File Systems

```{r}
#| label: crop_hook
#| echo: false
knitr::knit_hooks[["set"]](crop = knitr::hook_pdfcrop)
```

```{r}
#| label: setup
library(terra)
```

## Reading the data from a GeoTIFF file

- `vsicurl`: https://gdal.org/drivers/raster/vsicurl.html is a GDAL virtual file system that allows reading files from a URL.

```{r}
#| label: vsicurl
#| crop: true

r <- terra::rast(
  "/vsicurl/https://gebco2023.s3.valeria.science/gebco_2023_sub_ice_topo_cog.tif"
)

r

plot(r)
```

## vsigzip

- `vsigzip`: https://gdal.org/drivers/raster/vsigzip.html is a GDAL virtual file system that allows reading files from a gzip compressed file.

```{r}
#| label: vsigzip
#| crop: true

sic <- rast(
  "/vsigzip//vsicurl/ftp://ftp-projects.cen.uni-hamburg.de/seaice/AMSR2/3.125km/Arc_20201010_res3.125_pyres.nc.gz",
  "sea_ice_concentration"
)

sic

plot(sic)
```

From the documentation, we expect that data to be in [polar stereographic projection with a specific extent](https://nsidc.org/data/user-resources/help-center/guide-nsidcs-polar-stereographic-projection). However, the data does not have the correct extent and projection. We can set the extent and projection manually:

```{r}
#| label: projection
# Set the extent
ext(sic) <- ext(-3850000L, 3750000L, -5350000L, 5850000L)

# Set the polar stereographic projection
crs(sic) <- "EPSG:3413"

sic
```

Now the data has the correct extent and projection. Each pixel represents the sea ice concentration in a 3.125 km x 3.125 km area.

```{r}
#| crop: true

plot(sic)
```
